A Comparsion of Traffic Engineering Methods for Application Perceived Metrics

### 2015-07-21 tepaper/section1.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

Traffic engineering (TE) techniques are used by Internet Service Providers (ISPs) for computing network routing based on expected traffic demands. Our work focus on two aspects of ISP traffic engineering that have received less attention in prior work and are of key importance in a content-dominated network. First, the impact of traffic engineering on user-perceived peformance such as TCP file download times. Second, the impact of application adaptation on traffic engineering. In particular we focus on a form of adaptation enabled by \emph{location diversity}, or the ability of applications to download content from multiple locations. Location diversity is prevalent in the Internet due to several applications and services such as content delivery networks, peer-to-peer applications and mirrored websites.

While ISP traffic engineering schemes are commonly evaluated using link-utilization based metrics, these metrics are ill-suited to compare schemes based on user-perceived performance or network capacity, i.e., the factor of surge in traffic demand that a network can tolerate, under location diversity. As discussed previously in Chapter \ref{ch:te-background}, link utilization-based metrics may not reflect application performance that depends on other metrics such as propagtion delay, access link capacity and backbone link capacity also. Further, link utilization based-metrics, in particular MLU, do not reflect network capacity under location diversity as explained before.

@@

We show following results from comparing SPF achieved by TE schemes. The optimal TE scheme with perfect knowledge of traffic matrix and sub-optimal TE schemes like demand-aware TE by optimizing OSPF link weights \cite{fortz2000internet} end-up achieving the same SPF. Even the demand-oblivious scheme, \invcap, achieves an SPF at most 30\% worse than optimal TE. Overall, these results show that application adaptation to location diversity blurs differences between traffic engineering schemes.

The rest of the chapter is as follows. Section \ref{sec:locdiv-background} introduces location diversity and proposes a new capacity metric SPF to compare traffic engineering schemes. Section \ref{sec:exp_setup} presents our simulation setup. Section \ref{sec:app_performance} compares the application performance of TE schemes and Section \ref{sec:capacity} compares their achieved capacity under location diversity.

>>>>

Traffic engineering (TE) techniques are used by Internet Service Providers (ISPs) for computing network routing based on expected traffic demands \cite{rexford}. Our work focus on two aspects of ISP traffic engineering that have received less attention in prior work and are of key importance in a content-dominated network. The first is the impact of traffic engineering on user-perceived peformance such as TCP file download times. The second is the impact of application adaptation on traffic engineering. In particular we focus on a form of adaptation enabled by \emph{location diversity}, or the ability of applications to download content from multiple locations. Location diversity is prevalent in the Internet due to several applications and services such as content delivery networks, peer-to-peer applications and mirrored websites \cite{akamai-overview,bittorrent}.

While ISP traffic engineering schemes are commonly evaluated using link-utilization based metrics, these metrics are ill-suited to compare schemes based on user-perceived performance or network capacity, i.e., the factor of surge in traffic demand that a network can tolerate, under location diversity. As discussed previously in Chapter \ref{ch:te-background}, link utilization-based metrics may not reflect application performance that depends on other metrics such as propagtion delay, access link capacity and backbone link capacity also. Further, link utilization based-metrics, in particular MLU, do not reflect network capacity under location diversity.

@@

The explanation for the above results is that link loss rates and queuing delays remain negligibly small at typical Internet loads, and in fact until the utilization starts approaching the capacity. This observation above also been confirmed by explicit measurements on Internet backbones \cite{ExpRouterBuffer}, and is consistent with studies on ISP backbones showing that over 90\% of all packet loss is caused by interdomain routing fluctuations as opposed to high utilization \cite{SprintStudy} and 90\% of TCP flows experience no packet loss \cite{SprintBackbone}. Furthermore, end-to-end Internet path delays are known to be largely determined by propagation delays as opposed to queueing delays \cite{SprintBackbone,SingleHopDelay}.

Our comparison of the SPF of TE schemes in which application leverage location diversity by downloading content in parallel from all locations shows following somewhat surprising results. A small degree of location diversity (2-4 locations) 
increases SPF achieved by TE schemes by up to 2$\times$. But, the capacity increase benefits sub-optimal TE schemes, e.g., demand-aware TE via OSPF link weight optimization  \cite{fortz2000internet}, much more than the optimal TE scheme. As a result, all demand-aware TE schemes end-up achieving the same SPF as the optimal TE scheme. Even the demand-oblivious scheme, \invcap, achieves an SPF at most 30\% worse than optimal TE.

We also experiment with a second adaptation scheme in which users download content from a single location with the smallest propagation delay. 
In these experiments, an increase in location diversity yields little improvement in SPF of any TE scheme. This adaptation reduces the SPF of \opt\ to bridge the gap between \opt\ and sub-optimal schemes such as \optwt.
Overall, these results show that application adaptation to location diversity signficantly reduces the differences between TE schemes.

The rest of the chapter is organized as follows. Section \ref{sec:locdiv-background} introduces location diversity and proposes a new capacity metric SPF to compare traffic engineering schemes. Section \ref{sec:exp_setup} presents our simulation setup. Section \ref{sec:app_performance} compares the application performance of TE schemes and Section \ref{sec:capacity} compares their achieved capacity under location diversity.

### 2015-07-21 tepaper/section2.tex
several edits (tue jul 21)
updated graphs
added new citations
double spaced

This section introduces location diversity and propose a new metric to quantify the capacity achieved by traffic engineering schemes with location diversity.

>>>>

This section introduces location diversity and proposes a new metric to quantify the capacity achieved by traffic engineering schemes with location diversity.
